Reproducing Analyses for Phenotypic evolution is less constrained in complex

### 2018-10-08 analyses/reproduce_analyses.Rmd
update of current analyses

## Load required libraries ----
library(car)        # for homogeneity of variance test
library(tidyverse)  # for managing data
library(cowplot)    # pretty default ggplots
library(broom)      # for tidying multiple linear models
library(visreg)     # for quick visualizations of model effects
library(viridis)    # for color palette
library(lme4)       # for generalized linear mixed models
library(piecewiseSEM) # for calculating model R-square
library(latex2exp)  # using Latex notation for axes labels
library(kableExtra)
library(stringr)

@@

  # subset data for analysis
  filter(phenology == "early", Location == "tree",
         platy > 0 | ectos > 0 | pupa > 0) %>%                  # eliminate unknown sources of mortality
  filter(Treatment.focus == "Ectoparasitoid exclusion" & ectos < 1 | Treatment.focus == "Control") %>% # excluding any larva that were parasitized by an ectoparasitoid in the exclusion experiment.
  mutate(Foodweb = ifelse(Treatment.focus=="Control","Complex","Simple"),
         gall_survival = as.numeric(ifelse(pupa > 0, 1, 0)),
         egg_parasitoid = as.numeric(ifelse(platy > 0, 1, 0)),
         sc.Diam = as.numeric(scale(Gall_Height_mm)), 
         sc.log.Clutch = as.numeric(scale(log(gall_individuals))), 
         sc.log1p.Pref = as.numeric(scale(log1p(Density_per_100_shoots))),
         sc.DiamPoly1 = poly(sc.Diam,2)[,1],
         sc.DiamPoly2 = poly(sc.Diam,2)[,2],
         sc.log.ClutchPoly1 = poly(sc.log.Clutch,2)[,1],
         sc.log.ClutchPoly2 = poly(sc.log.Clutch,2)[,2],
         sc.log1p.PrefPoly1 = poly(sc.log1p.Pref,2)[,1],
         sc.log1p.PrefPoly2 = poly(sc.log1p.Pref,2)[,2])

@@

## Compute the matrix describing the curvature of the fitness landscape
curvature_fitness.landscape <- function(gammas, # matrix of nonlinear and correlational selection gradients 
                                        betas){ # vector of directional selection gradients
  gammas - betas %*% t(betas)
}

@@

@@

@@

```{r Boot Adjusted Model 2 for accurate slope estimates}
boot_model_2_adj <- bootMer(model_2_adj, FUN = fixef, nsim=n_boots_analysis, parallel="multicore", ncpus=32)

@@

```{r Fit Model 4}
model_4 <- update(model_3, .~. -Foodweb:sc.Diam -Foodweb:sc.log1p.Pref)
```

@@

As we did with model 2, we refit the model by removing the slope term so that we can fit different slopes for each interaction term.
```{r Fit Model 4 adjusted for estimating clutch slopes}
model_4_adj <- update(model_4, .~. -sc.log.Clutch)
```

@@

```{r Plot effect of Egg vs. Larval Parasitoid Model 2, echo=FALSE}
tidy_egg.larval_model_2_adj$term_plot <- factor(tidy_egg.larval_model_2_adj$term,
                            levels=c("I(sc.Diam^2)","I(sc.log.Clutch^2)","FoodwebEgg:I(sc.log1p.Pref^2)","FoodwebLarval:I(sc.log1p.Pref^2)","sc.Diam:sc.log.Clutch","sc.Diam:sc.log1p.Pref","sc.log.Clutch:sc.log1p.Pref"),
                            ordered = TRUE)

@@

```{r Fit Egg vs. Larval Parasitoid Model 4}
egg.larval_model_4 <- update(model_4, data=egg.vs.larval_df)
```

@@

```{r Fit Egg vs. Larval Parasitoid adjusted Model 4}
egg.larval_model_4_adj <- update(model_4_adj, data=egg.vs.larval_df)
```

@@

```{r Redo selection gradients}
bias.adj_selection_gradients <- mutate(selection_gradients, adjusted = "No") %>%
  # add bias adjusted directional selection gradient for Diameter in complex food-web
  add_row(term="sc.Diam", 
          gradient = complex_mean_brackets * (filter(tidy_common_coefs, term=="sc.Diam")$estimate - fixef(biased.diam_model_4)["sc.Diam"]) / complex_mean_fitness, 
          `2.5%` = complex_mean_brackets * (filter(tidy_common_coefs, term=="sc.Diam")$conf.low - fixef(biased.diam_model_4)["sc.Diam"]) / complex_mean_fitness, 
          `97.5%` = complex_mean_brackets * (filter(tidy_common_coefs, term=="sc.Diam")$conf.high - fixef(biased.diam_model_4)["sc.Diam"]) / complex_mean_fitness, 
          `Food web` = "Complex", 
          term_ord = "Diam", 
          adjusted="Yes") %>%
  # add bias adjusted directional selection gradient for Diameter in simple food-web
  add_row(term="sc.Diam", 
          gradient = simple_mean_brackets * (filter(tidy_common_coefs, term=="sc.Diam")$estimate - fixef(biased.diam_model_4)["sc.Diam"]) / simple_mean_fitness, 
          `2.5%` = simple_mean_brackets * (filter(tidy_common_coefs, term=="sc.Diam")$conf.low - fixef(biased.diam_model_4)["sc.Diam"]) / simple_mean_fitness, 
          `97.5%` = simple_mean_brackets * (filter(tidy_common_coefs, term=="sc.Diam")$conf.high - fixef(biased.diam_model_4)["sc.Diam"]) / simple_mean_fitness, 
          `Food web` = "Simple", 
          term_ord = "Diam", 
          adjusted="Yes")
```

@@

Appears to be selection for Platygaster to attack larva at high densities in more complex food webs (and a decrease in the variance). This may be a result of the poorer searching ability of parasitoids at very high gall densities, potentially due to saturation. This suggests that the fitness landscape may be dynamic because these selective effects (assuming there is heritable variation in the egg parasitoid) will change the following year. This could be an important discussion point to take into account for future work. This is different from simply altering the strength of selection, which will occur as species move along the fitness landscape, but this result suggests that the nature of the fitness landscape may actually be changing. For example, larval parasitoids could be driving selection on Platygaster to be very efficient at foraging, which will dampen once this pressure is removed, and then could also dampen the selection acting on the galling herbivore.
```{r Selection on Platygaster Traits, include=FALSE, eval=FALSE}
platy_selection_df <- gall_selection.df %>%
  filter(egg_parasitoid > 0) 
platy_selection_df$Foodweb <- relevel(factor(platy_selection_df$Foodweb), ref = "Simple") # this represents the baseline or absence of selection by larval parasitoids

platy_Diam <- lmer(sc.Diam ~ Foodweb + (1|Genotype/Plant_Position/Gall_Number), platy_selection_df)
summary(platy_Diam)
Anova(platy_Diam, test.statistic = "F")

platy_Clutch <- lmer(sc.log.Clutch ~ Foodweb + (1|Genotype/Plant_Position), platy_selection_df)
summary(platy_Clutch)
Anova(platy_Clutch, test.statistic = "F")

platy_Pref <- lmer(sc.log1p.Pref ~ Foodweb + (1|Genotype), platy_selection_df)
summary(platy_Pref)
Anova(platy_Pref, test.statistic = "F")

## Test whether food-web complexity influences nonlinear selection with Levene's test for homogeneity of variance between groups
# Note that this result should be interepreted cautiously since we cannot account for non-independence due to plant genotype, plant ID, or gall ID
leveneTest(sc.Diam ~ Foodweb, platy_selection_df)
leveneTest(sc.log.Clutch ~ Foodweb, platy_selection_df)
leveneTest(sc.log1p.Pref ~ Foodweb, platy_selection_df)

ggplot(platy_selection_df, aes(x=Foodweb, y=sc.Diam)) +
  stat_summary(fun="mean_cl_boot", geom="errorbar", color="red", width=0.1) +
  stat_summary(fun.y="mean", geom="point", color="black", size=3)

ggplot(platy_selection_df, aes(x=Foodweb, y=sc.log.Clutch)) +
  stat_summary(fun="mean_cl_boot", geom="errorbar", color="red", width=0.1) +
  stat_summary(fun.y="mean", geom="point", color="black", size=3)

ggplot(platy_selection_df, aes(x=Foodweb, y=sc.log1p.Pref)) +
  stat_summary(fun="mean_cl_boot", geom="errorbar", color="red", width=0.1) +
  stat_summary(fun.y="mean", geom="point", color="black", size=3)
```

```{r Slope and Curvature of Fitness Landscape, include=FALSE, eval=FALSE}
betas <- c(-0.5,-0.5,0.5)
gammas_matrix <- matrix(c(-0.1, 0.0, 0.0,
                          0.0, -0.1, 0.0,
                          0.0, 0.0, -0.1), 
                        nrow=3, byrow=T)

## G-matrix ----
# (1) Asssume equal additive genetic variance for all three traits
# (2) Assume no additive genetic covariance
# The first assumption may not be realistic, but violations of this assumption will not qualitatively alter our inferences about whether additive genetic variance will increase or decrease. Qualitative changes in additive genetic variance is only determined by the curvature of the fitness landscape. Violations of this assumption will only affect the rate of phenotypic evolution and magnitude of change in genetic variance.
# The second assumption may also not be realistic; however, it will not affect our results because we found no evidence of correlational selection (therefore, there can be no change in genetic covariance).
G_matrix <- matrix(c(1.0, 1.0, 1.0,
                     1.0, 1.0, 1.0,
                     1.0, 1.0, 1.0),
                   nrow=3, byrow=T)

# Equation 1 from Phillips and Arnold 1989, Evolution
delta_z <- G_matrix %*% betas
data.frame(trait = c("diam","clutch","pref"),
           delta_z = delta_z)

# Equation 2 from Phillips and Arnold 1989, Evolution
delta_G <- G_matrix * (gammas_matrix - betas %*% t(betas)) * G_matrix
delta_G
```

```{r LATEX NEED TO CHECK, eval=FALSE, include=FALSE}
$$\text{Slope}_{complex} = \beta_{complex} = \begin{pmatrix} \beta_{diam} \\ \beta_{clutch} \\ \beta_{pref} \end{pmatrix} = \begin{pmatrix} 0.4 \\ 0 \\ -0.3 \end{pmatrix}$$
$$\text{Slope}_{simple} = \beta_{simple} = \begin{pmatrix} \beta_{diam} \\ \beta_{clutch} \\ \beta_{pref} \end{pmatrix} = \begin{pmatrix} 0.3 \\ -0.2 \\ -0.3 \end{pmatrix}$$
$$\gamma_{complex} = \begin{pmatrix} \gamma_{diam}&& \\ \gamma_{diam,clutch}&\gamma_{clutch}& \\ \gamma_{diam,pref} & \gamma_{clutch,pref} &\gamma_{pref} \end{pmatrix} = \begin{pmatrix} 0.1&& \\ 0&0& \\ 0&0&-0.3 \end{pmatrix}$$
$$\gamma_{simple} = \begin{pmatrix} \gamma_{diam}&& \\ \gamma_{diam,clutch}&\gamma_{clutch}& \\ \gamma_{diam,pref} & \gamma_{clutch,pref} &\gamma_{pref} \end{pmatrix} = \begin{pmatrix} 0.1&& \\ 0&0& \\ 0&0&0 \end{pmatrix}$$
$$\text{Curvature} = \gamma - \beta \beta^T$$
$$\text{Curvature}_{complex} = \begin{pmatrix} 0.1&& \\ 0&0& \\ 0&0&0 \end{pmatrix}$$
$$\text{Curvature}_{simple} = \begin{pmatrix} 0.1&& \\ 0&0& \\ 0&0&0 \end{pmatrix}$$
```

```{r Effects of platy vs. ecto on absolute fitness, cache=TRUE, include=FALSE, eval=FALSE}

## Fit GLMM

platyecto_glmer <- glmer(gall_survival ~ Foodweb*(sc.Diam*sc.log.Clutch*sc.log1p.Pref +
                         I(sc.Diam^2) + I(sc.log.Clutch^2) + I(sc.log1p.Pref^2)) +
                         (1|Genotype/Plant_Position/Gall_Number),
                       data = filter(gall_selection.df, Foodweb == "Simple" | Foodweb == "Complex" & platy < 1), # exclude cases of egg-parasitism from Complex food web
                       family = binomial(link = logit), control=glmerControl(optimizer = "bobyqa"))

## Test for overdispersion
overdisp_fun(platyecto_glmer) # no evidence

## Compare statistical summaries of each model
summary(platyecto_glmer)
piecewiseSEM::rsquared(platyecto_glmer)

## Parse out mechanisms of food-web effects ----
# Use drop1 to test highest order effects and visreg to visualize the effects

# Marginal evidence that larval parasitoids are the drivers of nonlinear selection acting on female preference (disruptive selection)
drop1(platyecto_glmer, test="Chisq")
summary(platyecto_glmer)
visreg::visreg(platyecto_glmer, xvar="sc.log1p.Pref", by="Foodweb", scale = "response")

# Marginal evidence of nonlinear selection acting on gall size. Marginal evidence of correlational selection acting on all traits.
platyecto_glmer_2 <- update(platyecto_glmer, .~. -Foodweb:I(sc.Diam^2) -Foodweb:I(sc.log.Clutch^2) -Foodweb:I(sc.log1p.Pref^2) -Foodweb:sc.Diam:sc.log.Clutch:sc.log1p.Pref)
drop1(platyecto_glmer_2, test="Chisq") 
visreg::visreg(platyecto_glmer_2, xvar="sc.Diam", scale = "response") 
# plot 3-way interaction?

# Evidence of food-web treatment altering directional selection on clutch size
platyecto_glmer_3 <- update(platyecto_glmer_2, .~. -I(sc.Diam^2) -I(sc.log.Clutch^2) -I(sc.log1p.Pref^2) -sc.Diam:sc.log.Clutch:sc.log1p.Pref -Foodweb:sc.Diam:sc.log.Clutch -Foodweb:sc.Diam:sc.log1p.Pref -Foodweb:sc.log.Clutch:sc.log1p.Pref)
drop1(platyecto_glmer_3, test="Chisq") 
summary(platyecto_glmer_3)
visreg::visreg(platyecto_glmer_3, xvar="sc.log.Clutch", by="Foodweb", scale="response") # conflicting selection have net null effects in complex food web

# No evidence of clear fitness effects of platy vs. ectos on absolute fitness; evidence of directional selection acting on gall size; no evidence of directional selection on clutch size (in complex food web); evidence of directional selection acting on female preference
platyecto_glmer_4 <- update(platyecto_glmer_3, .~. -sc.Diam:sc.log.Clutch -sc.Diam:sc.log1p.Pref -sc.log.Clutch:sc.log1p.Pref -Foodweb:sc.Diam -Foodweb:sc.log.Clutch -Foodweb:sc.log1p.Pref)
drop1(platyecto_glmer_4, test="Chisq") 
summary(platyecto_glmer_4)
visreg::visreg(platyecto_glmer_4, xvar="Foodweb", scale="response")
visreg::visreg(platyecto_glmer_4, xvar="sc.Diam", scale="response")
visreg::visreg(platyecto_glmer_4, xvar="sc.log.Clutch", scale="response")
visreg::visreg(platyecto_glmer_4, xvar="sc.log1p.Pref", scale="response")
```

```{r For platy ecto comparison, include=FALSE, eval=FALSE}
## Estimate mean fitness and mean brackets (ref: Janzen and Stern 1998)
complex_predict <- predict(platyecto_glmer, newdata=filter(gall_selection.df, Foodweb == "Complex" & platy < 1), type="response")
complex_mean_brackets <- mean(complex_predict * (1 - complex_predict))
complex_mean_fitness <- mean(complex_predict)

simple_predict <- predict(platyecto_glmer, newdata=filter(gall_selection.df, Foodweb=="Simple"), type="response")
simple_mean_brackets <- mean(simple_predict * (1 - simple_predict))
simple_mean_fitness <- mean(simple_predict)

# test whether using alternative models influences estimes of mean "brackets" and mean fitness 
alt_predict <- predict(platyecto_glmer_2, newdata=filter(gall_selection.df, Foodweb == "Complex" & platy < 1), type="response")
alt_mean_brackets <- mean(alt_predict * (1 - alt_predict))
alt_mean_fitness <- mean(alt_predict)
complex_mean_brackets - alt_mean_brackets
complex_mean_fitness - alt_mean_fitness

platyecto_tidy_4 <- filter(tidy(platyecto_glmer_4, conf.int=T), term %in% c("(Intercept)","FoodwebSimple"))
platyecto_tidy_3 <- filter(tidy(platyecto_glmer_3, conf.int=T), 
                         term %in% c("sc.Diam","sc.log.Clutch","sc.log1p.Pref",
                                     "FoodwebSimple:sc.Diam","FoodwebSimple:sc.log.Clutch","FoodwebSimple:sc.log1p.Pref"))
platyecto_tidy_2 <- filter(tidy(platyecto_glmer_2, conf.int=T), 
                         term %in% c("sc.Diam:sc.log.Clutch","sc.Diam:sc.log1p.Pref","sc.log.Clutch:sc.log1p.Pref",
                                     "FoodwebSimple:sc.Diam:sc.log.Clutch","FoodwebSimple:sc.Diam:sc.log1p.Pref","FoodwebSimple:sc.log.Clutch:sc.log1p.Pref"))
platyecto_tidy <- filter(tidy(platyecto_glmer, conf.int=T), 
                       term %in% c("I(sc.Diam^2)","I(sc.log.Clutch^2)","I(sc.log1p.Pref^2)",
                                   "FoodwebSimple:I(sc.Diam^2)","FoodwebSimple:I(sc.log.Clutch^2)","FoodwebSimple:I(sc.log1p.Pref^2)",
                                   "sc.Diam:sc.log.Clutch:sc.log1p.Pref",
                                   "FoodwebSimple:sc.Diam:sc.log.Clutch:sc.log1p.Pref"))
## These alphas aren't the appropriate comparison, this is because when there is no interaction term, then the average is taken across all of the food webs. This is okay for testing the statistical significance, but not appropriate when I want to actually compare the magnitudes of the alphas and betas
#platyecto_tidy_4 <- filter(tidy(platyecto_glmer_4), term %in% c("(Intercept)","FoodwebSimple","sc.Diam","sc.log.Clutch","sc.log1p.Pref"))
#platyecto_tidy_3 <- filter(tidy(platyecto_glmer_3), term %in% c("sc.Diam:sc.log.Clutch","sc.Diam:sc.log1p.Pref","sc.log.Clutch:sc.log1p.Pref","FoodwebSimple:sc.Diam","FoodwebSimple:sc.log.Clutch","FoodwebSimple:sc.log1p.Pref"))
#platyecto_tidy_2 <- filter(tidy(platyecto_glmer_2), term %in% c("I(sc.Diam^2)","I(sc.log.Clutch^2)","I(sc.log1p.Pref^2)","sc.Diam:sc.log.Clutch:sc.log1p.Pref","FoodwebSimple:sc.Diam:sc.log.Clutch","FoodwebSimple:sc.Diam:sc.log1p.Pref","FoodwebSimple:sc.log.Clutch:sc.log1p.Pref"))
#platyecto_tidy <- filter(tidy(platyecto_glmer), term %in% c("FoodwebSimple:I(sc.Diam^2)","FoodwebSimple:I(sc.log.Clutch^2)","FoodwebSimple:I(sc.log1p.Pref^2)","FoodwebSimple:sc.Diam:sc.log.Clutch:sc.log1p.Pref"))

get_alphas <- bind_rows(platyecto_tidy_4, platyecto_tidy_3, platyecto_tidy_2, platyecto_tidy) %>% 
  mutate(alpha = round(estimate, 3), alpha_2.5 = round(conf.low, 3), alpha_97.5 = round(conf.high, 3), P = round(p.value, 3)) %>%
  select(term, alpha, alpha_2.5, alpha_97.5, P)

complex_alphas <- filter(get_alphas, !grepl("FoodwebSimple", term))
simple_alphas <- filter(get_alphas, grepl("FoodwebSimple", term))

# remember that the selection gradients need to be multiplied by 2 for quadratic gradients.
coefs_gradients <- data.frame(Variable = complex_alphas$term, 
           a_complex = complex_alphas$alpha,
           a_complex_2.5 = complex_alphas$alpha_2.5,
           a_complex_97.5 = complex_alphas$alpha_97.5,
           a_simple = complex_alphas$alpha + simple_alphas$alpha,        # the baseline is the coefficient in the complex food web, therefore we simply add the corresponding coefficients from the simple food web to get the coefficients on the same scale as the complex web
           a_simple_2.5 = complex_alphas$alpha + simple_alphas$alpha_2.5, # same addition process for confidence intervals
           a_simple_97.5 = complex_alphas$alpha + simple_alphas$alpha_97.5) %>% 
  mutate(b_complex = round(complex_mean_brackets * a_complex / complex_mean_fitness, 3),
         b_complex_2.5 = round(complex_mean_brackets * a_complex_2.5 / complex_mean_fitness, 3),
         b_complex_97.5 = round(complex_mean_brackets * a_complex_97.5 / complex_mean_fitness, 3),
         b_simple = round(simple_mean_brackets * a_simple / simple_mean_fitness, 3),
         b_simple_2.5 = round(simple_mean_brackets * a_simple_2.5 / simple_mean_fitness, 3),
         b_simple_97.5 = round(simple_mean_brackets * a_simple_97.5 / simple_mean_fitness, 3)) %>%
  filter(Variable != "(Intercept)")

select(coefs_gradients, Variable, a_complex:a_simple_97.5)
select(coefs_gradients, Variable, b_complex:b_simple_97.5)

bind_rows(mutate(select(coefs_gradients, Variable, a=a_complex, a_2.5=a_complex_2.5, a_97.5=a_complex_97.5), Foodweb = "Complex"),
          mutate(select(coefs_gradients, Variable, a=a_simple, a_2.5=a_simple_2.5, a_97.5=a_simple_97.5), Foodweb = "Simple")) %>%
  ggplot(., aes(x = Variable, fill=Foodweb, color=Foodweb)) +
  geom_point(aes(y=a), position = position_dodge(width=0.5)) +
  geom_errorbar(aes(ymax = a_97.5, ymin = a_2.5), position = position_dodge(width=0.5)) +
  coord_flip() +
  geom_hline(yintercept=0, linetype="dotted")

bind_rows(mutate(select(coefs_gradients, Variable, b=b_complex, b_2.5=b_complex_2.5, b_97.5=b_complex_97.5), Foodweb = "Complex"),
          mutate(select(coefs_gradients, Variable, b=b_simple, b_2.5=b_simple_2.5, b_97.5=b_simple_97.5), Foodweb = "Simple")) %>%
  ggplot(., aes(x = Variable, fill=Foodweb, color=Foodweb)) +
  geom_point(aes(y=b), position = position_dodge(width=0.5)) +
  geom_errorbar(aes(ymax = b_97.5, ymin = b_2.5), position = position_dodge(width=0.5)) +
  coord_flip() +
  geom_hline(yintercept=0, linetype="dotted")

inverse_logit(complex_alphas$alpha[1]) - complex_mean_fitness # small difference in mean fitness estimates compared to intercept in complex food web
inverse_logit(complex_alphas$alpha[1] + simple_alphas$alpha[1]) - simple_mean_fitness # larger difference in mean fitness estimates from intercept in simple food web, unclear why...perhaps because the mean phenotypes don't match up as nicely?
```

```{r include=FALSE, eval=FALSE}
## Calculating same coefficients as for simple vs. complex food web

## Estimate mean fitness and mean brackets (ref: Janzen and Stern 1998)
ecto_predict <- predict(platyecto_glmer, newdata=filter(gall_selection.df, Foodweb=="Complex" & platy < 1), type="response")
ecto_mean_brackets <- mean(ecto_predict * (1 - ecto_predict))
ecto_mean_fitness <- mean(ecto_predict)

platy_predict <- predict(platyecto_glmer, newdata=filter(gall_selection.df, Foodweb=="Simple"), type="response")
platy_mean_brackets <- mean(platy_predict * (1 - platy_predict))
platy_mean_fitness <- mean(platy_predict)

>>>>

## Load required libraries ----
library(evolqg) # need to load before dplyr
library(car)        # for homogeneity of variance test
library(tidyverse)  # for managing data
library(cowplot)    # pretty default ggplots
library(broom)      # for tidying multiple linear models
library(visreg)     # for quick visualizations of model effects
library(viridis)    # for color palette
library(lme4)       # for generalized linear mixed models
library(piecewiseSEM) # for calculating model R-square
library(latex2exp)  # using Latex notation for axes labels
library(kableExtra)
library(stringr)
library(MVN)        # assessing multivariate normality assumption

@@

  # subset data for analysis
  filter(phenology == "early", Location == "tree",
         platy > 0 | ectos > 0 | pupa > 0) %>%                  # eliminate unknown sources of mortality
  filter(Treatment.focus == "Ectoparasitoid exclusion" & ectos < 1 | Treatment.focus == "Control") %>% # excluding any larva that were parasitized by an ectoparasitoid in the exclusion experiment.
  mutate(Foodweb = ifelse(Treatment.focus=="Control","Complex","Simple"),
         gall_survival = as.numeric(ifelse(pupa > 0, 1, 0)),
         egg_parasitoid = as.numeric(ifelse(platy > 0, 1, 0)),
         sc.Diam = as.numeric(scale(Gall_Height_mm)), 
         sc.log.Clutch = as.numeric(scale(log(gall_individuals))), 
         sc.log1p.Pref = as.numeric(scale(log1p(Density_per_100_shoots))))

@@

@@

## Predicting evolutionary dynamics 
evolutionary_dynamics <- function (betas,    # one-column matrix of directional selection gradients
                                   gammas,   # matrix of nonlinear selection gradients
                                   G_0,      # initial G-matrix
                                   z_0,      # one-column matrix of initial trait values
                                   t) {       # evolutionary time steps 
    # Equation 2 from Phillips and Arnold 1989, Evolution
    G <- rep(list(G_0), t) 
    for (i in 1:t) G[[i + 1]] <- {
      G[[i]] + G[[i]] %*% (gammas - betas %*% t(betas)) %*% G[[i]] # may need to add mutation/recombination matrix (see Phillips and Arnold 1989)
    }

    # Equation 1 from Phillips and Arnold 1989, Evolution
    z <- rep(list(z_0), t) 
    for (i in 1:t) z[[i + 1]] <- {
        z[[i]] + G[[i]] %*% betas
    }
    return(list(z=z,G=G))
}

## Number of evolutionary time steps for simulation
n_evo_sims <- 200 # 1000

## Tidy up evolutionary simulations
tidy_evo_sims <- function(evo_sims) {
  tidy_sim <- data.frame(matrix(unlist(evo_sims$z), ncol=3, byrow=T)) %>% 
    transmute(Time=1:length(evo_sims$z), Diam=round(X1,2), Clutch=round(X2,2), Pref=round(X3,2)) %>%
    filter(Diam < 1, Diam > -1, Clutch < 1, Clutch > -1, Pref < 1, Pref > -1)

  return(tidy_sim)
}

@@

# Evaluating assumption of multivariate normality

We used graphical checks to evaluate whether our transformations of traits resulted in a multivariate normal distribution. As you can see from the multivariate qq-plot, it seems reasonable that our data approximates a multivariate normal distribution. Of course, statistical tests suggest otherwise. But as an example of how silly these tests can be, look at the distribution of gall diameter. It is the closest ecological variable I've ever seen that is normally distributed, yet a Shapiro-Wilks test says otherwise.

```{r MVN Assumption, echo=FALSE}

mvn_assumption <- mvn(select(gall_selection.df, sc.Diam, sc.log.Clutch, sc.log1p.Pref), multivariatePlot = "qq", univariatePlot = "histogram")

# mvn(select(gall_selection.df, sc.Diam, sc.log.Clutch, sc.log1p.Pref), multivariatePlot = "qq")

mvn_assumption$univariateNormality[1, ]
```

@@

```{r Boot Adjusted Model 2 for accurate slope estimates, cache=TRUE}
boot_model_2_adj <- bootMer(model_2_adj, FUN = fixef, nsim=n_boots_analysis, parallel="multicore", ncpus=32)

@@

```{r Fit Model 4, cache=TRUE}
model_4 <- update(model_3, .~. -Foodweb:sc.Diam -Foodweb:sc.log1p.Pref)
```

@@

As we did with model 2, we refit the model by removing the slope term so that we can fit different slopes for each interaction term. 
```{r Fit Model 4 adjusted for estimating clutch slopes, cache=TRUE}
model_4_adj <- update(model_4, .~. -sc.log.Clutch)
```

@@

```{r Plot effect of Egg vs. Larval Parasitoid adjusted Model 2, echo=FALSE}
tidy_egg.larval_model_2_adj$term_plot <- factor(tidy_egg.larval_model_2_adj$term,
                            levels=c("I(sc.Diam^2)","I(sc.log.Clutch^2)","FoodwebEgg:I(sc.log1p.Pref^2)","FoodwebLarval:I(sc.log1p.Pref^2)","sc.Diam:sc.log.Clutch","sc.Diam:sc.log1p.Pref","sc.log.Clutch:sc.log1p.Pref"),
                            ordered = TRUE)

@@

```{r Fit Egg vs. Larval Parasitoid Model 4, cache=TRUE}
egg.larval_model_4 <- update(model_4, data=egg.vs.larval_df)
```

@@

```{r Fit Egg vs. Larval Parasitoid adjusted Model 4, cache=TRUE}
egg.larval_model_4_adj <- update(model_4_adj, data=egg.vs.larval_df)
```

@@

```{r Redo selection gradients, include=FALSE}
bias.adj_selection_gradients <- mutate(selection_gradients, adjusted = "No") %>%
  # add bias adjusted directional selection gradient for Diameter in complex food-web
  add_row(term="sc.Diam", 
          gradient = complex_mean_brackets * (filter(tidy_common_coefs, term=="sc.Diam")$estimate - fixef(biased.diam_model_4)["sc.Diam"]) / complex_mean_fitness, 
          `2.5%` = complex_mean_brackets * (filter(tidy_common_coefs, term=="sc.Diam")$conf.low - fixef(biased.diam_model_4)["sc.Diam"]) / complex_mean_fitness, 
          `97.5%` = complex_mean_brackets * (filter(tidy_common_coefs, term=="sc.Diam")$conf.high - fixef(biased.diam_model_4)["sc.Diam"]) / complex_mean_fitness, 
          `Food web` = "Complex", 
          term_ord = "Diam", 
          adjusted="Yes") %>%
  # add bias adjusted directional selection gradient for Diameter in simple food-web
  add_row(term="sc.Diam", 
          gradient = simple_mean_brackets * (filter(tidy_common_coefs, term=="sc.Diam")$estimate - fixef(biased.diam_model_4)["sc.Diam"]) / simple_mean_fitness, 
          `2.5%` = simple_mean_brackets * (filter(tidy_common_coefs, term=="sc.Diam")$conf.low - fixef(biased.diam_model_4)["sc.Diam"]) / simple_mean_fitness, 
          `97.5%` = simple_mean_brackets * (filter(tidy_common_coefs, term=="sc.Diam")$conf.high - fixef(biased.diam_model_4)["sc.Diam"]) / simple_mean_fitness, 
          `Food web` = "Simple", 
          term_ord = "Diam", 
          adjusted="Yes")
```

@@

Appears to be selection for Platygaster to attack larva at high densities in more complex food webs (and a decrease in the variance). This may be a result of the poorer searching ability of parasitoids at very high gall densities, potentially due to saturation. This suggests that the fitness landscape may be dynamic because these selective effects (assuming there is heritable variation in the egg parasitoid) will change the following year. This could be an important discussion point to take into account for future work. This is different from simply altering the strength of selection, which will occur as species move along the fitness landscape, but this result suggests that the nature of the fitness landscape may actually be changing. For example, larval parasitoids could be driving selection on Platygaster to be very efficient at foraging, which will dampen once this pressure is removed, and then could also dampen the selection acting on the galling herbivore.
```{r Selection on Platygaster Traits, cache=TRUE}

## Subset the data to analyse change in gall phenotypes associated with Platygaster
platy_selection_df <- gall_selection.df %>%
  filter(egg_parasitoid > 0) 
platy_selection_df$Foodweb <- relevel(factor(platy_selection_df$Foodweb), ref = "Simple") # this represents the baseline or absence of selection by larval parasitoids

## Linear mixed models of linear selection differentials (change in mean phenotype)
platy_Diam <- lmer(sc.Diam ~ Foodweb + (1|Genotype/Plant_Position/Gall_Number), platy_selection_df)

platy_Clutch <- lmer(sc.log.Clutch ~ Foodweb + (1|Genotype/Plant_Position), platy_selection_df)

platy_Pref <- lmer(sc.log1p.Pref ~ Foodweb + (1|Genotype), platy_selection_df)

## Levene's test for quadratic selection differentials (change in phenotypic variance)
# Note that this result should be interepreted cautiously since we cannot account for non-independence due to plant genotype, plant ID, or gall ID
platy_Diam_levene <- leveneTest(sc.Diam ~ Foodweb, platy_selection_df)
platy_Clutch_levene <- leveneTest(sc.log.Clutch ~ Foodweb, platy_selection_df)
platy_Pref_levene <- leveneTest(sc.log1p.Pref ~ Foodweb, platy_selection_df)
```

```{r Plot models of Platygaster selection differentials, echo=FALSE}
visreg_platy_Diam <- visreg(platy_Diam, xvar="Foodweb", gg=TRUE) + ylab("Chamber diameter (SD)") + xlab("")
visreg_platy_Clutch <- visreg(platy_Clutch, xvar="Foodweb", gg=TRUE) + ylab("Clutch size (SD)") + xlab("")
visreg_platy_Pref <- visreg(platy_Pref, xvar="Foodweb", gg=TRUE) + ylab("Female preference (SD)") + xlab("")

plot_grid(visreg_platy_Diam, visreg_platy_Clutch, visreg_platy_Pref, align="hv", ncol=3)
```

```{r Calculate Selection Differentials for Platygaster, cache=TRUE, include=FALSE}

## Calculation linear selection differentials
boot_platy_Diam <- bootMer(platy_Diam, FUN = fixef, nsim=n_boots_analysis, parallel="multicore", ncpus=32)
tidy_platy_Diam <- tidy(boot_platy_Diam, conf.int=T) %>% filter(term=="FoodwebComplex") %>% mutate(term="Diam")

boot_platy_Clutch <- bootMer(platy_Clutch, FUN = fixef, nsim=n_boots_analysis, parallel="multicore", ncpus=32)
tidy_platy_Clutch <- tidy(boot_platy_Clutch, conf.int=T) %>% filter(term=="FoodwebComplex") %>% mutate(term="Clutch")

boot_platy_Pref <- bootMer(platy_Pref, FUN = fixef, nsim=n_boots_analysis, parallel="multicore", ncpus=32)
tidy_platy_Pref <- tidy(boot_platy_Pref, conf.int=T) %>% filter(term=="FoodwebComplex") %>% mutate(term="Pref")

platy_linear_selection <- bind_rows(tidy_platy_Diam, tidy_platy_Clutch, tidy_platy_Pref) %>% select(Term=term, Estimate=statistic, `2.5%`=conf.low, `97.5%`=conf.high)
platy_linear_selection$Significance <- c("NS","NS","P < 0.05")

## Calculate quadratic selection differentials
# statistical significance is based on Levene's test of homogeneity of variances
platy_quadratic_selection <- platy_selection_df %>%
  group_by(Foodweb) %>%
  summarise_at(c("sc.Diam","sc.log.Clutch","sc.log1p.Pref"), var) %>%
  gather(key = Term, value = Variance, -Foodweb) %>%
  spread(Foodweb, Variance) %>%
  mutate(Estimate = Complex - Simple) %>%
  select(Term, Estimate)
platy_quadratic_selection$Term <- c("Diam^2","Pref^2","Clutch^2")
platy_quadratic_selection$Significance <- c("NS","P < 0.05","NS")
```

```{r Table of Platygaster selection differentials, echo=FALSE}
platy_selection_differentials <- bind_rows(platy_linear_selection, platy_quadratic_selection)

knitr::kable(platy_selection_differentials, caption="Table of selection differentials acting on the egg parasitoid Platygaster")
```
